---
title: "Practical session: ex1.Rmd"
author: '  '
date: '   '
output:
  pdf_document: default
  html_document:
    df_print: paged
---


During this practical session, we run the examples included in this file and may test the same code for different data sets. 

## Creating data reports

```{r, , echo = TRUE, eval=FALSE}

dat <- iris
dat[1:3,1] <- dat[3:7,2] <- dat[8:10,5] <- NA

# interesting also, optional investigation of the package at:
# https://cran.r-project.org/web/packages/missMethods/ 

# is.na(dat)[1:5,]
# mean(dat[1,], na.rm=TRUE) 
# dat[!complete.cases(dat),]
# newdat <- na.omit(dat)

## two methods of creating automatic reports
## to be run separately, line by line:
## outputs are html

DataExplorer::create_report(data=dat) 
             ##   , output_format = html_document()) 

dataReporter::makeDataReport(data=dat, output = "html", replace = TRUE)

```

Note: remember to check your browser and your working directory!

## Exploration of missing values

_Reference: https://epirhandbook.com/en/missing-data.html#assess-missingness-in-a-data-frame_

```{r, echo=TRUE, eval=TRUE}

dat <- iris
dat[1:3,1] <- dat[3:7,2] <- dat[8:10,5] <- NA
head(dat)

# percentage of missing values
naniar::pct_miss(dat)
# Percent of rows with missing values
naniar::pct_miss_case(dat)     # use n_complete() for counts
# Percent of complete rows (no values missing)  
naniar::pct_complete_case(dat) # use n_complete() for counts

# visualisation of missing values
library(ggplot2)

mice::md.pattern(dat)

naniar::gg_miss_var(dat, show_pct = TRUE, facet="Species")

naniar::gg_miss_fct(dat, Sepal.Length)

ggplot(
  data = dat,
  mapping = aes(x = Petal.Length, y = Sepal.Length)) +     
  naniar::geom_miss_point()

# create indicators of missing values
library(magrittr)
shadowed_dat <- dat %>% 
  naniar::bind_shadow()
# compare distributions of variable X1 between groups defined by 
# missing and non-missing values of variable X2
ggplot(data = shadowed_dat,               # data frame with shadow columns
  mapping = aes(x = Petal.Length,         # numeric or date column
                colour = Sepal.Length_NA)) + # shadow column of interest
  geom_density()                          # plots the density curves

```



## Testing the mechansim of missing data

_Reference: Little, Roderick J. A. 1988. "A Test of Missing Completely at Random for Multivariate Data with Missing Values." Journal of the American Statistical Association 83 (404): 1198--1202. doi: 10.1080/01621459.1988.10478722_ , which:

  -  uses a multivariate test statistic for association between 
missing status of some variable values and observed values of other variables.

  - becomes a simple t-test if the data set has:
only two variables and one of them has missing values.


```{r , echo = TRUE, eval=FALSE}
## use test for MCAR, as shown at:
# https://naniar.njtierney.com/reference/mcar_test.html

naniar::mcar_test(airquality) # from base R

naniar::mcar_test(dat)   # our example

```



