---
title: "Practical session: ex2.Rmd"
author: '  '
date: '   '
output:
  pdf_document: default
  html_document:
    df_print: paged
---

## Compare imputation methods


### Select and add new methods and apply the method described in the course slides (see pages 42-46) and addapted from https://stefvanbuuren.name/fimd/sec-evaluation.html 


#### Function: create data 

```{r, , echo = TRUE, eval=TRUE}

create.data <- function(beta = 1, sigma2 = 1, n = 50,
                        run = 1) {
  set.seed(seed = run)
  x <- rnorm(n)
  y <- beta * x + rnorm(n, sd = sqrt(sigma2))
  cbind(x = x, y = y)
}
```


#### Function: create missing values

```{r, , echo = TRUE, eval=TRUE}

make.missing <- function(data, p = 0.5){
  rx <- rbinom(nrow(data), 1, p)
  data[rx == 0, "x"] <- NA
  data
}

```



### Function: imputations with variable methods

See: https://cran.r-project.org/web/packages/mice/mice.pdf 

```{r, , echo = TRUE, eval=TRUE}

test.impute <- function(data, m , method = "method", ...) {
  imp <- mice(data, method = method, m = m, print = FALSE, ...)
  fit <- with(imp, lm(y ~ x))
  tab <- summary(pool(fit), "all", conf.int = TRUE)
  as.numeric(tab[2, c("estimate", "2.5 %", "97.5 %")])  
  ## 2 indicates the line of interest in the result
}

```


### Simulate results for new methods

```{r, , echo = TRUE, eval=TRUE, warning=FALSE}

## should increase the number of simulations!
library(mice)

simulate <- function(runs = runs) {
  res <- array(NA, dim = c(5, runs, 3))
  dimnames(res) <- list(c("norm.predict", "norm.nob", "mean", "norm" , "lasso.norm")
                       , as.character(1:runs),
                        c("estimate", "2.5 %","97.5 %"))
  for(run in 1:runs) {
    data <- create.data(run = run)
    data <- make.missing(data)
    
    ## select several imputation methods  
    res[1, run, ] <- test.impute(data, method = "norm.predict", m = 5)
    res[2, run, ] <- test.impute(data, method = "norm.nob", m=5)
    res[3, run, ] <- test.impute(data, method = "mean", m=5)
    res[4, run, ] <- test.impute(data, method = "norm", m=5)
    res[5, run, ] <- test.impute(data, method = "lasso.norm", m=5)
    
  }
  res}

res <- simulate(runs=20)
apply(res, c(1, 3), mean, na.rm = TRUE)

```




### Evaluate performance of new methods

```{r, , echo = TRUE, eval=TRUE}

true <- 1
RB <- rowMeans(res[,, "estimate"]) - true
PB <- 100 * abs((rowMeans(res[,, "estimate"]) - true)/ true)
CR <- rowMeans(res[,, "2.5 %"] < true & true < res[,, "97.5 %"])
AW <- rowMeans(res[,, "97.5 %"] - res[,, "2.5 %"])
RMSE <- sqrt(rowMeans((res[,, "estimate"] - true)^2))
data.frame(RB, PB, CR, AW, RMSE)

```





### Resources https://cran.r-project.org/web/packages/mice/mice.pdf, 
see all methods at page 76 of this paper.


## Compare data sets with _daff_ package

### Resource: exercise addapted from the book on "Statistical data cleaning with applications in R"(2018) by Mark van der Loo

```{r, , echo = TRUE, eval=TRUE, warning=FALSE}
library(validate)
data(retailers, package="validate")
library(simputation)
library(daff)

retailers2a <- simputation::impute_lm(retailers, turnover ~ total.rev)
# or, when grouping by company size: # ***
retailers2 <- simputation::impute_lm(retailers, turnover ~ total.rev| size)
# compute diff ('retailers' is the reference set)
d <- diff_data(retailers, retailers2)
# get the diff as a data.frame
d$get_data()

# change name of one column and check what happens
retailers3 <- dplyr::rename(retailers2, income = turnover)
d1 <- diff_data(retailers, retailers3)
head(d1$get_data())

retailers3_reconstructed <- patch_data(retailers, d)
all.equal(retailers3_reconstructed, retailers3)
## [1] "Names: 1 string mismatch"

## Complement: about cell changes!
validate::cells(start = retailers, step1 = retailers2)

```



## More resources (optional) at:

https://www.ebpi.uzh.ch/dam/jcr:dc0cef17-29c7-4e61-8d33-e690561ab7ae/mi_intro20191001.pdf

https://bookdown.org/mike/data_analysis/imputation-missing-data.html#mice-multivariate-imputation-via-chained-equations

https://stefvanbuuren.name/fimd/ch-multilevel.html

https://epirhandbook.com/en/missing-data.html#assess-missingness-in-a-data-frame






