---
title: "Practical session: ex3.Rmd"
author: '  '
date: '   '
output:
  pdf_document: default
  html_document:
    df_print: paged
---


During this practical session, we run the examples included in this file and may test the same code for different data sets. 


## Analysis of longitudinal and or clustered data

### Examples of analysis for data with complex correlation structures

- Data types: longitudinal, spatial, repeated, clustered, multivariate measurements 

- Multilevel models: $P(Y|b)=F(b)$ and $P(b|b1)=F1(b1)$ and $P(b1)...$


#### Create simple clustered data

```{r, echo = TRUE, eval=TRUE, warning=FALSE}

set.seed(1234)  # this will allow you to exactly duplicate your result
Ngroups = 100
NperGroup = 3
N = Ngroups * NperGroup
groups = factor(rep(1:Ngroups, each = NperGroup))

u = rnorm(Ngroups, sd = .5)
e = rnorm(N, sd = .25)
x = rnorm(N)
y = 2 + .5 * x + u[groups] + e

d = data.frame(x, y, groups)

```



#### Model and performance

```{r,  echo = TRUE, eval=TRUE, warning=FALSE, message=FALSE}
library(lme4)
library(lmerTest)

model0 = lmer(y ~ 1 + (1|groups), data=d) #, REML=FALSE)
model = lmer(y ~ x + (1|groups), data=d) #, REML=FALSE)

summary(model)

confint(model)

VarCorr(model)

library(ggplot2)
ggplot(aes(x, y), data=d) +
  geom_point()

re =lme4::ranef(model)$groups
ggplot2::qplot(x = re[[1]], geom = 'density', xlim = c(-3, 3))

performance::icc(model)

performance::r2(model, by_group = TRUE)


#F-test with Satterthwaite adjusted degrees of freedom:
anova(model) 

# likelihood ratio tests (Deviance Difference Test) by single term deletion 
lmerTest::ranova(model)

```



#### Plots for model checking

```{r,  echo = TRUE, eval=TRUE, warning=FALSE, fig.align="left"}

plot(fitted(model), resid(model, type = "pearson")) 
abline(0,0, col="red")

qqnorm(resid(model))
qqline(resid(model), col = "red") 

qqnorm(ranef(model)$groups[,1] )
qqline(ranef(model)$groups[,1], col = "red")

library(merTools)
library(ggplot2)

fastdisp(model)    ### displaying the model coeff.est and coeff.se

#fixed effects  
feEx <- FEsim(model, 1000)
plotFEsim(feEx) +
  theme_bw() + labs(title = "Coefficient Plot",
                    y = "Median Effect Estimate", x = "")

#random effects
reEx <- REsim(model)
head(reEx)

##lattice::dotplot(ranef(model, condVar=TRUE))

p1 <- plotREsim(reEx)
p1

```


 
### Bayesian style evaluation of model

 
```{r,  echo = TRUE, eval=TRUE, warning=FALSE}

### Bayesian style credible intervals
library(arm)
sim_fit <- arm::sim( model, 5000)  
bayestestR::hdi(sim_fit)
bayestestR::hdi(sim_fit, effects="fixed")  # or "all"   # or "random"
## too long
# bayestestR::hdi(sim_fit, effects="random") 
bayestestR::eti(sim_fit)

xf <- bayestestR::ci(sim_fit, ci = c(.5, .8, .95), effects = "fixed")
plot(xf)

## too big
# xr <- bayestestR::ci(sim_fit, ci = c(.5, .8, .95), effects = "random") 
# plot(xr)

```



#### Comparing models

```{r,  echo = TRUE, eval=TRUE, warning=FALSE}

table_build_models_1 <- 
   sjPlot::tab_model(
     
  model0, 
  model,
     
     show.ci = FALSE,
     show.reflvl = TRUE,
     show.se = TRUE,
     auto.label = FALSE,
     string.se = "SE",
     show.icc = TRUE,
     show.aic = TRUE, 
     digits=3,
     dv.labels = c("model0", "model")
    ## ,file="comparingModels1.doc"
     
   )
 
 ## table_build_models_1
 
```




### Imputations for multilevel models

See: https://stefvanbuuren.name/fimd/ch-multilevel.html


#### Data: structured and with missing values

```{r,  echo = TRUE, eval=TRUE, warning=FALSE}

library(mice)
data("brandsma", package = "mice")

#di <- brandsma[, c("sch", "lpo", "sex", "den")]

di <- brandsma[, c("sch", "lpo")]


# School number: cluster variable
# Language test post: outcome at pupil level
# Sex of pupil: predictor at pupil level
# School denomination: predictor at school level

mice::md.pattern(di, plot = TRUE)

```



#### Imputation methods

```{r,  echo = TRUE, eval=TRUE, warning=FALSE}

# sample: Find imputations by random sampling from the observed values in lpo. 
#         This method ignores sch;
# pmm: Single-level predictive mean matching with the school indicator coded as
#         a dummy variable;
# 2l.pan: Multilevel method using the linear mixed model to draw univariate imputations;
# 2l.norm: Multilevel method using the linear mixed model with heterogeneous error variances;
# 2l.pmm: Predictive mean matching based on predictions from the linear mixed model, 
#         with random draws from the regression coefficients and the random effects, using five donors.

library(miceadds)
library(pan)
methods <- c("sample", "pmm", "2l.pan", "2l.norm", "2l.pmm")
result <- vector("list", length(methods))
names(result) <- methods
for (meth in methods) {
  di <- brandsma[, c("sch", "lpo")]
  pred <- make.predictorMatrix(di)
  pred["lpo", "sch"] <- -2  ## says that "sch" is the grouping variable
  # maxit = 1 since only one variable has missing values, no need to iterate
  result[[meth]] <- mice(di, pred = pred, meth = meth,
                         m = 10, maxit = 1,
                         print = FALSE, seed = 82828)
}


## look at results
## for example, method 3 (but you may choose any other!)

# for method 3, the imputed missing values are
d3imp <- result[[3]]$imp
str(d3imp)
#head(d3imp[[2]][,1:5], 5)

# for method 3, select one complete data set, for example set 9
d3 <- complete(result[[3]],9)
head(d3)
# or another set: method 3, set 2
d3b <- complete(result[[3]],2)

# fit MLM models
m300 <- lm(lpo~1, data=di)
m30 <- lmer(lpo ~ 1 + (1|sch), data=di, REML=FALSE)
m3 <- lmer(lpo ~ 1 + (1|sch), data=d3, REML=FALSE)
m3b <- lmer(lpo ~ 1 + (1|sch), data=d3b, REML=FALSE)


# Compare models
table_build_models <- 
   sjPlot::tab_model(
     
   m300,
   m30,
   m3,
   m3b,
     
     show.ci = FALSE,
     show.reflvl = TRUE,
     show.se = TRUE,
     auto.label = FALSE,
     string.se = "SE",
     show.icc = TRUE,
     show.aic = TRUE, 
     digits=3,
     dv.labels = c("m300", "m30", "m3", "m3b")
    ## ,file="comparingModels.doc"
     
   )
 
## table_build_models  ## in viewer
  
  
```


## Time series analysis

### Resource: Robert Hyndman's book online, at https://otexts.com/fpp2/ 


```{r, echo = TRUE, eval=TRUE, warning=FALSE}

# time series decomposition
library(forecast)
data(elecequip, package = "fpp")
fit <- stl(elecequip, t.window=13, s.window="periodic",
  robust=TRUE)
# forecasting with decomposition
fit %>% seasadj() %>% naive() %>%
  autoplot() + ylab("New orders index") +
  ggtitle("Naive forecasts of seasonally adjusted data")

# now forecast() function is applied to the "stl" object. We need to specify the method being used on the seasonally adjusted data, and the function will do the re-seasonalising for you. 
fit %>% forecast(method="naive") %>%
  autoplot() + ylab("New orders index")

# using neural networks
# data(sunspotarea, package = "fpp")
# or the data used previously
fitNN <- nnetar(elecequip, lambda=0)
autoplot(forecast(fitNN,h=30))
# with prediction intervals, also:
autoplot(forecast(fitNN,h=30, PI=TRUE))

```



## Interpretation of Machine Learning results


```{r,  echo = TRUE, eval=TRUE, warning=FALSE}
## see the iml vignette page
set.seed(42)
library("iml")
library("randomForest")

# median house values from Boston neighbourhoods
data("Boston", package = "MASS")
head(Boston)
rf <- randomForest(medv ~ ., data = Boston, ntree = 50)

X <- Boston[which(names(Boston) != "medv")]
predictor <- Predictor$new(rf, data = X, y = Boston$medv)

imp <- FeatureImp$new(predictor, loss = "mae")
library("ggplot2")
plot(imp)
imp$results

interact <- Interaction$new(predictor)
plot(interact)

interact1 <- Interaction$new(predictor, feature = "crim")
plot(interact1)

effs <- FeatureEffects$new(predictor)
plot(effs)

tree <- TreeSurrogate$new(predictor, maxdepth = 2)
plot(tree)

head(tree$predict(Boston))

```


### More resources at:

https://cran.r-project.org/web/packages/iml/vignettes/intro.html

https://cran.r-project.org/web/packages/DALEX/index.html

https://christophm.github.io/interpretable-ml-book/agnostic.html



```{r , echo = TRUE, eval=TRUE,}

sessionInfo()

```


